---
title: "Lab4"
author: "Lily Bates"
date: "2023-02-14"
output: html_document
---

## 1 Load and check data

1. Load tidyverse library. Set the chunk option message=FALSE to avoid the loading messages to
show up in the document. (See https://faculty.washington.edu/otoomet/info201-book/
r-markdown.html#rmarkdown-syntax-chunk-options).

```{r message=FALSE}
library(tidyverse)
```


2. Load the dataset. Remember that the working directory of the knitting process is the same
as the folder where your rmarkdown file is located As above, ensure that the informative
messages are not displayed in the document.

```{r message=FALSE}
getwd()

IncomeHeight <- read_delim("income-height.csv.bz2")
```



3. How many observations and variables do you have in this dataset? Answer in a complete
sentence like “This dataset contains 4 rows and 2 columns” using the inline code chunks.

```{r}
rows <- nrow(IncomeHeight)
columns <- ncol(IncomeHeight)
```

This dataset contains `r rows` rows and `r columns` columns


4. List the variable names of the dataset, separated by comma, using a complete sentence.
Hint: check out function str_flatten_comma() in stringr library.

```{r message=FALSE}
VarNames <- str_to_sentence(names(IncomeHeight))
str_flatten_comma(VarNames)
```

The variable names are: `r VarNames` 
Is there anyway to get rid of vector code ?? 

5. How many missings are there in variables income, height, age and sex? Again, answer in a
complete sentence, injecting the numbers into the text like There are 5 missings in variable
“age”, 4 missings in ....

```{r}
missingIncome <- IncomeHeight %>% 
  filter(is.na(income)) %>% 
  count()

missingHeight <- IncomeHeight %>% 
  filter(is.na(height)) %>% 
  count()

missingAge <- IncomeHeight %>% 
  filter(is.na(age)) %>% 
  count()

missingSex <- IncomeHeight %>% 
  filter(is.na(sex)) %>% 
  count()
```

There are `r missingIncome` missing in variable "income", `r missingHeight` 
missing in variable "height", `r missingAge` missing in variable "age", and 
`r missingSex` missing in variable "sex"

_Double checked, there really are no missings in these variables_

## 2 Do taller people earn more?
1. Find the average height in data. Answer in a complete sentence, round the result as appro-
priate.

```{r}
aveHeight <- IncomeHeight %>% 
  summarise(mean(height)) %>% 
  round(1)
```

The average height in the dataset is `r aveHeight` inches tall. 

2. Compute the average income for those who are shorter than the average height and those who
are taller than the average height. Do this in a code chunk and show the code and the output.
Comment your finding in text below.
Hint: you can use group_by(), grouping by a logical condition like group_by(height >
avgHeight).

```{r}
diffIncome <- IncomeHeight %>% 
  group_by(height > 67.1) %>% ## aveHeight variable not working???
  summarise(mean(income)) %>% 
  round(0)

diffIncome

```
Short people on average make 30,848 which is much less than the tall person average of 53,721.

3. But does the difference also hold if we do it separately by sex? Now compute average income
separately for men and women who are shorter and taller than the corresponding male and
female average.
Hint: you may first group by sex, then compute for every person if they are taller than the
average, and then group again, this time by sex and if they are taller. Finally use summarize()
to compute the average income.

```{r}
femShort <- IncomeHeight %>% 
  filter(sex == "female") %>% 
  mutate(tall = height > mean(height)) %>% 
  filter(tall == FALSE) %>% 
  summarise(mean(income)) %>% 
  round(0) %>% 
  prettyNum()

maleShort <- IncomeHeight %>% 
  filter(sex == "male") %>% 
  mutate(tall = height > mean(height)) %>% 
  filter(tall == FALSE) %>% 
  summarise(mean(income)) %>% 
  round(0) %>% 
  prettyNum()

femTall <- IncomeHeight %>% 
  filter(sex == "female") %>% 
  mutate(tall = height > mean(height)) %>% 
  filter(tall == TRUE) %>% 
  summarise(mean(income)) %>% 
  round(0) %>% 
  prettyNum()

maleTall <- IncomeHeight %>% 
  filter(sex == "male") %>% 
  mutate(tall = height > mean(height)) %>% 
  filter(tall == TRUE) %>% 
  summarise(mean(income)) %>% 
  round(0) %>% 
  prettyNum()
```

4. Now comment your results in a sentence–something like “short women earn $27,000 while
tall women earn $32,000 in average...”. However, the numbers should be extracted from your
results above using inline code chunks, not just copied in the text.
Do not forget to round the numbers as appropriate! You may also check out the function
prettyNum for better printing of large numbers.

Short women earn on average `r femShort`, while tall women earn `r femTall`. 
Short men earn on average `r maleShort`, while tall men earn `r maleTall`.

5. Make a scatterplot where you plot height versus income. Do not forget to add informative
labels. Comment the plot in text.
Hint: if using base-R plotting then you can use pch="." to make the scatterplot of dots, not
circles.

```{r}
dfPlot <- IncomeHeight %>% 
  plot(income~height, data=. , 
       xlab= "Height in Inches", ylab = "Income",
       pch=".")
```

Scatter plot of height vs income `r dfPlot`